---
title: "CMSC320 Final - League of Legends Analysis"
author: "Harish Kumar"
date: "May 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## The Rise of E-Sports

For millenia, humans have embraced the adrenaline and thrill of competition; hunting, gathering, sport, and even war, are all testament to this. Today, civilization continues to glorify competition, as can be seen in global events (Olympics, Commonwealth)




## The Dataset



## Grabbing the Data

We will turn the data into a data table that R can use for analysis. A *data table* is a row by column representation of all the data. Each row represents an **entity**, which is a real-world object, person, or concept; in the case of our dataset, the entity is a single League of Legends game. Now, each column represents an **attribute**, which is a characteristic for an entity. In our case, a LoL game can have an id, a duration, a winner, some statistics (like First Blood), and more.

Our data is stored locally. We create a string that describes the computer location of the dataset, which has been stored as a CSV file. A **Comma-seperated values** (CSV) file is just that, a bunch of values seperated by columns. The `read_csv()` method of R, parses out the commas, analyzes the values, and creates a data table that is much easier for us to handle when we do analysis.

And since this is a huge dat table, we only want to look at some of the values as a preview; we do this with the `head()` function.

```{r cars}
url <- 'data/games.csv'
lol <- read_csv(url)
head(lol)
```


