---
title: "CMSC320 Final - League of Legends Analysis"
author: "Harish Kumar"
date: "May 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## The Rise of E-Sports

For millenia, humans have embraced the adrenaline and thrill of competition; hunting, gathering, sport, and even war, are all testament to this. Today, civilization continues to glorify competition, as can be seen in global events (Olympics, Commonwealth)




## The Dataset



## Grabbing the Data

We will turn the data into a data table that R can use for analysis. A *data table* is a row by column representation of all the data. Each row represents an **entity**, which is a real-world object, person, or concept; in the case of our dataset, the entity is a single League of Legends game. Now, each column represents an **attribute**, which is a characteristic for an entity. In our case, a LoL game can have an id, a duration, a winner, some statistics (like First Blood), and more.

Our data is stored locally. We create a string that describes the computer location of the dataset, which has been stored as a CSV file. A **Comma-seperated values** (CSV) file is just that, a bunch of values seperated by columns. The `read_csv()` method of R, parses out the commas, analyzes the values, and creates a data table that is much easier for us to handle when we do analysis.

And since this is a huge dat table, we only want to look at some of the values as a preview; we do this with the `head()` function.

```{r cars}
url <- 'data/games.csv'
lol <- read_csv(url)
head(lol)
```

## Observing and Managing the Data

We have loaded the dataset into a nice data table that we can use to do our analysis. Let us look at it; the dataset is 51490 rows by 61 columns! That means we have 51490 games, and 61 attributes/characteristics. Let's have a quick overview as to what they are:

  - `gameId`: the ID number of the game
  - `creationTime`: the time of when the game was created, in numeric format
  - `gameDuration`: how long the game went, in seconds
  - `seasonId`: all games are from Season 9
  - `winner`: which team won the game (Team 1 or 2)
  - `firstBlood`: which team got the first kill of the game
  - `firstTower`: which team destroys an enemy tower first
  - `firstInhibitor`: which team destroys the Inhibitor first
  
The remaining columns describe the champions (characters) that both teams used during the game.

Now, as data analysts we need to ask ourselves the question: Is the data tidy? That is, is the data stored in a format that is easily understandable for humans, easily processable for computers, and easily streamlined for analysis? To ensure this, we can follow **Tidy Data Principles**.

Envisioned by Hadley Wickham, the creator of the magnificent Tidyverse R package, the Tidy Data Principles promote three key features of rectangular data tables (like the ones we have).

  1. Each attribute forms a column
  2. Each entity forms a row
  3. Only one entity forms a table
  
Lucky for us, our dataset follows all three of these principles! Each row is a game; each column is a characteristic about the game, and the table contans only game entities!

But just because our data is clean, does not mean we are done! Let's see what else we can tidy up, to optimize our data analysis. First, it would be nice if we can see the creation time attribute in a more human-friendly date format. We will use the lubridate package to do this.

```{r dates}

```






























































































































































