---
title: "CMSC320-Final-Tutorial-Project"
author: "Jerry Qian & Harish Kumar"
date: "May 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
#Customize this to your directory!
setwd("D:/Documents/Semester4/CMSC320/CMSC320FinalProject")
```

# Raw data
```{r}
library(dplyr)
library(tidyr)
library(lubridate)

games <- read.csv("D:/Documents/Semester4/CMSC320/CMSC320FinalProject/data/games.csv") %>% 
  
head(games, 100)
```

##Exercise 1: Scatterplot
```{r}
library(ggplot2)

ggplot(gapminder, aes(x=year, y=lifeExp)) + geom_point() + geom_smooth(method="lm")
```
###Question 1: Is there a general trend (e.g., increasing or decreasing) for life expectancy across time? Is this trend linear? (answering this qualitatively from the plot, you will do a statistical analysis of this question shortly)
Yes, there is an increasing trend for life expectancy across time. I added a regression line to show this. It appears to be relatively linear except for a slight decrease in "acceleration" since 1990. This could potentially be logarithmic.

##Violin plot
```{r}
library(tidyverse)
library(ggplot2)

gapminder %>%
  ggplot(aes(x=factor(year), y=lifeExp)) +
    geom_violin() +
    labs(title="Life expectancy over time",
         x = "year",
         y = "life expectancy")
```


